What really matters (and to whom): Investigating heterogeneous effects in high-dimensional conjoint analyses

### 2024-12-04 Manuscript files/Manuscript.qmd
TR: Updates to Manuscript

Following the nested causal quantities structure outlines above, IMCEs are calculated by averaging the OMCEs for each individual.
Specifically, the OMCE estimates from step 2 are summed and divided by $J \times K$, that is, the number of profiles times the number of rounds (which is the number of total observations).

@@

The first tool uses random forests, an ensemble of decision trees, to investigate which covariates are the important ones in the prediction of differences in the IMCE distributions for all attribute levels (i.e., the effect heterogeneity).
Specifically, for each attribute level, a random forest is trained where the covariates $X_i$ are used as predictors.
Subsequently, variable importance measures (VIMPs) are derived to assess which covariates are driving the data partitioning. Briefly, the VIMPs are calculated by permuting the covariates (i.e., by introducing random noise to its values).
If predictions change noticeably, this is an indication that the covariate is important for predicting the outcome (i.e., IMCE distributions).
Of note, this is done separately for every combination of covariate and attribute level because the importance of a certain covariate may differ with respect to different attribute levels.
This results in a comprehensive VIMP-heatmap, which we show as an example in the results section.

@@

```{r attributes, echo=FALSE}
#| output: asis
#| label: tbl-attributes
library(kableExtra)

@@

@@

For AMCEs, theoretically most favorable events served as baseline. 
As can be seen, increased attribute levels (i.e., less favourable events) were usually linked to significantly decreased levels of support. 
For example, increasing the risk of a nuclear strike from the reference level 'Not present (0%)' to 'Moderate (10%)' decreased the average probability of a participant choosing the latter profile by about 8.5 percentage points. 
The AMCE point estimates were in line with the AMCEs reported in CITE, which is why we refrain from an in-depth discussion here and refer the interested reader to CITE.
HIER WAS DAZU SCHREIBEN DAS WIR HÖHERE SEs HABEN? DAZU SCHREIBEN DIE HALT IN IHREM PAPER EH GAR NIX

As a next step, we predicted the subject-level IMCEs using the subject-level covariates. 
Based on this random forest model, we then calculated variable importance measures which served as indicators for variables associated with effect heterogeneity. 
We found the highest variable importance values for the subject-level covariates country, gender, and the items asking participants about whether they generally support supply of weapons to countries defending themselves against other country's invasions (Q10_10) and whether Ukraine should make any concessions in order to achieve peace (Q15_7).

These variable importance measures do not provide any more interpretability than which covariates in general seem to matter, which is why we further substantiated the previous analysis by fitting single decision trees to different IMCEs. 
IMCEs for this were selected due to content-related purposes and to highlight how decision trees facilitate interpretation of effect heterogeneity. 
As can be seen in @image-panel (upper left panel), for participants from countries different to the UK the probability to select a given profile decreased by -12.5% if the respective profile was associated with 16,000 (vs. 4,000) Ukrainian civilian casualties. 
For participants from the UK, the larger number of civilian casualties was also linked to a lower profile selection probability, although, the attribute level was less influential for their decision. 
In general, we saw the UK vs other countries split in many different trees possibly indicating slight differences in attitudes towards weapon supplies across these countries. 
Moreover, whether participants were female (within both UK and Non-UK subgroups) predicted

Apart from this, participants' attitude about whether the Ukraine should make any concessions to Russia if this would lead to a peace agreement predicted some level of effect heterogeneity to an interesting extent. 
For example, with regard to Russian soldiers killed, people who responded that Ukraine should not make any concessions, on average were more likely to select profiles with a larger number of Russian military casualties (100,000 vs. 25,000). 
Interestingly, out of the investigated IMCE decision trees this was the only case for which we observed splits resulting in nodes with different signs meaning that the probability of profile selection decreased in one child node and increased in the other. 
These results have, however, to be interpreted with caution as the split cannot be interpreted as a sign of statistically significant differences. 
As for this result, though having different signs, the differences in probability are still small (less than 2 percentage points) and close to zero. 
Lastly, for the attribute of how much cession of Ukrainian territory a selected profile would result in, we once again found the general attitude towards Ukrainian concessions to predict effect heterogeneity. 
People who responded that Ukraine should make no concessions were 16 percentage points less likely to select profiles with the 'comparably moderate negative outcome' of Ukraine loosing 8% of their territory (resembling the 2014 Line of Conflict) (vs. reference level 'None (0%)'). 
For people that did not select the no concessions option, profile selection probabilities for the same attribute level decreased by only 2.4 percentage points.

@@

<!-- -Vergleich mit “normaler” Subgruppen Analyse aka vergleich mit deren paper -->
However, as already outlined in many other papers, these AMCEs cannot be used to reason about absolute profile selection probabilities but are relative in nature. 
Accordingly a given AMCE has always to be interpreted as change in profile selection probability *relative to a given reference level* of the same attribute. 
This still holds when comparing IMCEs across different subsets of participants, no matter whether these partitions were or are said-to-be defined theory-based or were generated by data (as in the previous approach). 
Accordingly the average IMCEs in the different terminal nodes of the decision trees always are relative to the baseline profile selection probability taking into account the reference level *and* the subgroups of participants. 
The often applied approach of calculating Marginal Means, on the other hand, provides marginalized profile selection probabilities for different attribute levels. 
This allows for straightforward reading off the absolute profile selection probabilities. 
However, as marginal means are solely descriptive, they do not allow for any causal interpretation. 
Accordingly, whether AMCEs or MMs are used should be guided by the kind of conclusions that shall be made. 
As in the present case, the goal was to understand how different subject-level covariates affect decisions-making among different weapon supply attributes, AMCEs provide the right choice for answering this question.

-Vergleich mit “normaler” Subgruppen Analyse --> aka vergleich mit deren paper

@@

Third, more simulated and empirical research on cjbart is needed to properly assess the stability of its solutions.
Internally, the *cjbart()* function as implemented by CITE calls the *mc-pbart()* function from the *BART* package CITE.
This function estimates complex relations using the BART model described above and uses a seed (which defaults to 99).
Thus, when repeatedly used on the same data, users should always receive the same results (of AMCEs).
Simulation studies should be conducted to assess the variability in results when seeds are changed.
Additional variance in results is then introduced when using random forests to assess variable importances.
This step should be included in these simulations to investigate whether, and if yes, how strongly changes in BART results are perpetuated in the variable importance results.
Other factors that might influence cjbarts results are the specific settings of hyperparameters.
Hyperparameters are second-order parameters that cannot be estimated by the data but that have to be set prior to estimating the model.
Examples of hyperparameters of the cjbart model are the number of individual trees that are grown sequentially and the number of cut points that are investigated (i.e., the number of "segments" of, for example, age for which a split is considered).
In our analyses, we used the default values of these hyperparameters (50 trees and 100 cut points).
However, it would also be possible to tune these values, that is, to try out different settings and choose the most optimal values (with respect to some evaluation metric).
Because tuning hyperparameters can be computationally expensive, it is an interesting topic of future methodological research to assess whether tuning in the case of cjbart is worth the additional computational burden.
Tuning hyperparameters would also be possible for the random forest when calculating variable importances.
However, random forests have been shown to have good performance with default values of its hyperparameters (CITE), so we believe the focus should be on the hyperparameters of the upstream BART model.

Forth and last, cjbart is not the only method to take potential heterogeneity in attribute level effects into account.
As mentioned in the introduction, two other methods are FactorHet by CITE and the CRT approach by CITE.
FactorHet works based on a Bayesian mixture of regularized logistic regression analyses.
The goal is to identify maximally heterogeneous groups of respondents for which similar patterns of attribute level effects apply.
The membership probability to belong to a certain group, also called cluster, is modeled as a function of the person-level covariates.
This allows users to investigate person characteristics (like country of origin or age) that are indicative of belonging to certain patterns of attribute level effects.
While this yields interpretable results with regard to the question of "what matters to whom", it has the important downside that the number of clusters is a hyperparameter that has to be set prior to estimating the model.
Thus, the question arises which model to select.
It might be a good strategy to estimate the model with different numbers of clusters and then use domain expertise and interpretability
to choose a solution.
However, this should be made with caution to not let post-hoc explanation guide the decision.
Future research concering FactorHet could develop model selection strategies, like information criteria, that are comparable across solutions with differing numbers of clusters and allow for a statistical decision.
The other method, CRT by CITE, is a statistical hypothesis testing approach based on conditional randomization tests.
The goal of this approach is to assess a specific attribute of interest influences participants' choice behavior in any way, given all the other attributes.
The advantage of CRT is that this answers the most basic question one can ask in conjoint experiments, while being assumption-free.
This is due to the fact that the underlying testing approach works solely on the randomization of attributes.
However, in these analyses, it remains challenging to investigate the influence of person-level covariates on individual-level preferences (i.e., IMCEs).
Taken together, all novel approaches have their own strengths and weaknesses when trying to unravel heterogeneity in attribute level effects.
As in every statistical analyses, the choice of methods and models crucially depends on the research question and desired inferences one wants to make.
Because our goal was to complement the existing analyses reported by CITE with a maximally exploratory approach, potentially informing future more confirmatory research, we found that cjbart best fits this purpose.

>>>>

Following the nested structure of causal quantities outlines above, IMCEs are calculated by averaging the OMCEs for each individual.
Specifically, the OMCE estimates from step 2 are summed and divided by $J \times K$, that is, the number of profiles times the number of rounds (which is the number of total observations).
<!-- Sollen wir (wie im Paper) noch dazuschreiben, dass man basierend auf den Posterior Draws der OMCEs auch Credibility Intervals basteln kann, dann würde vielleicht auch der Bayesian Aspekt nochmal betont werden -->

@@

The first tool uses random forests, an ensemble of decision trees, to investigate which covariates are the important ones in the prediction of differences in the IMCE distributions for all attribute levels (i.e., the effect heterogeneity).
Specifically, for each attribute level, a random forest is trained where the covariates $X_i$ are used as predictors.
Subsequently, variable importance measures (VIMPs) are derived to assess which covariates are driving the data partitioning. Briefly, the VIMPs are calculated by permuting the covariates (i.e., by introducing random noise to its values).
<!-- nit: is random noise introduced (in terms of: added) or more the complete information on the variable destroyed by permuting all the values in X -->
If predictions change noticeably, this is an indication that the covariate is important for predicting the outcome (i.e., IMCE distributions).
Of note, this is done separately for every combination of covariate and attribute level because the importance of a certain covariate may differ with respect to different attribute levels.
This results in a comprehensive VIMP-heatmap, which we show as an example in the results section.

@@

<!-- falls wir Platz sparen oder die Tabelle auf eine Seite bekommen wollen: Reference Category mit Asterisk oder so anzeichnen, dann haben wir nur 2 Spalten und weniger Zeilenumbrüche vermutlich -->
<!-- Sollen wir hier noch ihre theoretischen 4 Cluster kurz reinschreiben? Also im Text oder in der Tabelle? Diese Aufteilung in Human Costs, Costs of Aid, Strategic Risks, Containment of Aggression (resolve)? -->
```{r attributes, echo=FALSE}
#| output: asis
#| label: tbl-attributes
library(kableExtra)

@@

These AMCEs are supplemented with the AMCEs reported in the original preprint by CITE, presented in @fig-amces-preprint.

![Preprint AMCEs](figures/AMCEs_Original_Preprint.png){#fig-amces-preprint}

@@

For AMCEs, theoretically most favorable events served as baseline. 
As can be seen, increased attribute levels (i.e., less favourable events) were usually linked to significantly decreased levels of support. 
For example, increasing the risk of a nuclear strike from the reference level 'Not present (0%)' to 'Moderate (10%)' decreased the average probability of a participant choosing the latter profile by about 8.5 percentage points. 
The AMCE point estimates were in line with the AMCEs reported in CITE, which is why we refrain from an in-depth discussion here and refer the interested reader to CITE.
HIER WAS DAZU SCHREIBEN, DASS WIR HÖHERE SEs HABEN? DAZU SCHREIBEN DIE HALT IN IHREM PAPER EH GAR NIX

As a next step, we fit a random forest model to predict subject-level IMCEs using the subject-level covariates. 
Based on this model, we then calculated variable importance measures which served as indicators for variables associated with effect heterogeneity. 
<!-- Wie sollen wir hier auf die Items referenzieren? brauchen vermutlich Item Kurznamen -->
We found the highest variable importance values for the subject-level covariates country, gender, and the items asking participants about whether they generally support supply of weapons to countries defending themselves against other country's invasions (Q10_10) and whether Ukraine should make any concessions in order to achieve peace (Q15_7).

As already outlined, these variable importance measures do not provide any more interpretability than which covariates in general seem to matter.
This is why we further substantiated the previous analysis by fitting single decision trees to different IMCEs. 
IMCEs for this were selected based on contentual characteristics and to highlight how decision trees facilitate interpretation of effect heterogeneity. 
As can be seen in @image-panel (panel a), for participants from countries different to the UK the probability to select a given profile decreased by -12.5 percentage points if the respective profile was associated with 16,000 (vs. 4,000) Ukrainian civilian casualties. 
For participants from the UK, the larger number of civilian casualties was also linked to a lower profile selection probability, although, the attribute level was less influential for their decision. 
In general, many trees performed splits based on whether participants were from the UK or the other countries included. 
For example, this was also the case for the attribute regarding the risk of a nuclear strike when comparing the risk of '10%' with the reference level of '0%' (@image-panel (panel c).
These UK versus Non-UK splits possibly indicate slight differences in attitudes towards weapon supplies across these countries. 
Moreover, for female participants (within both UK and Non-UK subgroups) the probability for selecting profiles with 16,000 vs. 4,000 Ukrainian civilian casualties, decreased stronger than for male participants, although differences were less pronounced than for the country covariate.

Apart from this, participants' attitude about whether the Ukraine should make any concessions to Russia if this would lead to a peace agreement predicted some level of effect heterogeneity to an interesting extent. 
For example, with regard to Russian soldiers killed (@image-panel (panel b), people who responded that Ukraine should not make any concessions, had a higher probability to select profiles with a larger number of Russian military casualties (100,000 vs. 25,000). 
Interestingly, out of the investigated IMCE decision trees this was the only case for which we observed splits resulting in nodes with different signs meaning that the probability of profile selection decreased in one child node and increased in the other. 
<!-- @Philipp: hast du hier gut ne Quelle, mit der wir einfach sagen können dass die Bäume basierend auf Impurity splitten und das eben nicht signifikanz bedeutet -->
These results have, however, to be interpreted with caution as the split cannot be interpreted as a sign of statistically significant differences between the terminal nodes. 
Moreover, for this result, though having different signs, the differences in probability are still small (less than 2 percentage points) and close to zero. 
Lastly, for the attribute regarding how much cession of Ukrainian territory a selected profile would result in (@image-panel (panel d), we once again found the general attitude towards Ukrainian concessions to predict effect heterogeneity. 
People who responded that Ukraine should make no concessions were 16 percentage points less likely to select profiles with the 'comparably moderate negative outcome' of Ukraine loosing 8% of their territory (resembling the 2014 Line of Conflict) (vs. reference level 'None (0%)'). 
For people that did not select the no concessions option, profile selection probabilities for the same attribute level decreased by only 2.4 percentage points.

@@

<!-- -Vergleich mit “normaler” Subgruppen Analyse aka vergleich mit deren paper -->
However, as already outlined by many authors (CITE), thes above AMCEs cannot be used to reason about absolute profile selection probabilities but instead are relative in nature. 
Accordingly a given AMCE has always to be interpreted as change in profile selection probability *relative to a given reference level* of the same attribute. 
This still holds when comparing IMCEs across different subsets of participants, no matter whether these partitions were (or were said-to-be) defined theory-based (as in classical subgroup analysis) or generated in a more data-driven manner (as in the previous approach). 
Accordingly the average IMCEs in the different terminal nodes of the decision trees always are relative to the baseline profile selection probability taking into account the reference level *and* the subgroups of participants. 
Accordingly, in order to reason about the absolute selection probability for a given attribute level using the associated AMCE, the absolute selection probability for the respective reference level must be known. 
The often applied approach of calculating Marginal Means, provides exactly these marginalized absolute profile selection probabilities for different attribute levels. 
This allows for straightforward reading off the absolute profile selection probabilities. 
However, as marginal means are solely descriptive, they do not allow for any causal interpretation. 
Accordingly, whether AMCEs or Marginal Means are used should be guided by the kind of conclusions that shall be made. 
As in the present case, the goal was to understand how different subject-level covariates affect decisions-making in the presence of different weapon supply attributes, AMCEs provide the right choice for answering this question.

@@

Third, more simulated and empirical research on cjbart is needed to properly assess the stability of its solutions.
Internally, the *cjbart()* function as implemented by CITE calls the *mc-pbart()* function from the *BART* package CITE.
<!-- Nur auf Mac ... :D  -->
This function estimates complex relations using the BART model described above and uses a seed (which defaults to 99).
Thus, when repeatedly used on the same data, users should always receive the same results (of AMCEs).
Simulation studies should be conducted to assess the variability in results when seeds are changed.
Additional variance in results is then introduced when using random forests to assess variable importances.
This step should be included in these simulations to investigate whether, and if yes, how strongly changes in BART results are perpetuated in the variable importance results.
Other factors that might influence cjbarts results are the specific settings of hyperparameters.
Hyperparameters are second-order parameters that cannot be estimated by the data but that have to be set prior to estimating the model.
Examples of hyperparameters of the cjbart model are the number of individual trees that are grown sequentially and the number of cut points that are investigated (i.e., the number of "segments" of, for example, age for which a split is considered).
In our analyses, we used the default values of these hyperparameters (50 trees and 100 cut points).
However, it would also be possible to tune these values, that is, to try out different settings and choose the most optimal values (with respect to some evaluation metric).
Because tuning hyperparameters can be computationally expensive, it is an interesting topic of future methodological research to assess whether tuning in the case of cjbart is worth the additional computational burden.
Tuning hyperparameters would also be possible for the random forest when calculating variable importances.
However, random forests have been shown to have good performance with default values of its hyperparameters (CITE), so we believe the focus should be on the hyperparameters of the upstream BART model.

Forth and last, cjbart is not the only method to take potential heterogeneity in attribute level effects into account.
As mentioned in the introduction, two other methods are FactorHet by CITE and the CRT approach by CITE.
FactorHet works based on a Bayesian mixture of regularized logistic regression analyses.
The goal is to identify maximally heterogeneous groups of respondents for which similar patterns of attribute level effects apply.
The membership probability to belong to a certain group, also called cluster, is modeled as a function of the person-level covariates.
This allows users to investigate person characteristics (like country of origin or age) that are indicative of belonging to certain patterns of attribute level effects.
<!-- Sollen wir hier einfach die Plots noch irgendwie in einen Appendix packen? ist zwar nicht Paper-Style aber so als Bericht ja vielleicht nich schlecht? Nachdem wir uns die Arbeit ja gemacht haben, könnte man das ja vielleicht schon machen? -->
While this yields interpretable results with regard to the question of "what matters to whom", it has the important downside that the number of clusters is a hyperparameter that has to be set prior to estimating the model.
Thus, the question arises which model to select.
It might be a good strategy to estimate the model with different numbers of clusters and then use domain expertise and interpretability
to choose a solution.
However, this should be made with caution to not let post-hoc explanation guide the decision.
Future research concerning FactorHet could develop model selection strategies, like information criteria, that are comparable across solutions with differing numbers of clusters and allow for a statistical decision.
The other method, CRT by CITE, is a statistical hypothesis testing approach based on conditional randomization tests.
The goal of this approach is to assess whether a specific attribute of interest influences participants' choice behavior in any way, given all the other attributes.
The advantage of CRT is that this answers the most basic question one can ask in conjoint experiments, while being assumption-free.
This is due to the fact that the underlying testing approach works solely on the randomization of attributes.
However, in these analyses, it remains challenging to investigate the influence of person-level covariates on individual-level preferences (i.e., IMCEs).
Taken together, all novel approaches have their own strengths and weaknesses when trying to unravel heterogeneity in attribute level effects.
As in every statistical analysis, the choice of methods and models crucially depends on the research question and desired inferences one wants to make.
Because our goal was to complement the existing analyses reported by CITE with a maximally exploratory approach, potentially informing future more confirmatory research, we found that cjbart best fits this purpose.
